---
title: "GAE7"
author: "Jaslyn Stamey"
format: 
  html:
    theme: custom.scss
    embed-resources: true
    code-tools: true

csl: ecology.csl

editor: visual
bibliography: references.bib
---

```{r}
#| label: SetUp
#| message: FALSE
#| echo: FALSE

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
  ) 

library(emmeans)
library(lmerTest)
library(lattice)
library(effectsize)
library(car)
library(ggfortify)
library(patchwork)
library(broom)
library(latex2exp)       
library(kableExtra)
library(tidyverse)

options(knitr.kable.NA = '')
```

## Introduction

@Pintar.etal2018 conducted a field mesocosm experiment to explore the relative effects of predation risk and nutrient availability (a pdf of their publication is included in the GitHub repository). The predation risk manipulation was accomplished by adding either 0, 1, or 2, Golden Topminnows (*Fundulus chrysotus*) to each mesocosm. These small fish are surface-feeding predators and are known to elicit colonization and oviposition avoidance among dytiscid beetles and *Culex* mosquitoes, but not in some hydrophilid beetles. Nutrient-level treatments were created by the addition of 0, 4, or 8 g of rabbit chow (Small World Rabbit Food, Mannapro, St. Louis, MO, 40% protein).

## Methods

Please see @Pintar.etal2018 for greater detail on experimental design and data collection. In brief, the experiment was set up as a 3 X 3 factorial treatment design (9 treatment combinations). The experiment was replicated in 3 separate locations at 3 different times for a total of 9 replicates. This should have resulted in 81 observations total, however some fish died during the experiment rendering those observations not being included in the data resulting in only 69 observations.

```{r }
#| label: dataImport

df1 <- read_csv("Pintar_etal_2018.csv",
                show_col_types = FALSE)

# Block = 1 to 9 for the 9 spatiotemporal blocks (3 locations X 3 times)
# Pool = 1 to 9 for the plastic wading pools used for the experiment
# Location = 1 to 3 for the 3 fields where the pools were deployed
# Time = a to c for the 3 time periods used for each 2-week run of the exp.
# Nutrient = 1 to 3 for the 3 nutrient levels
# Fish = 1 to 3 for the 3 fish densities
# Culex_Egg_Rafts = number of egg rafts collected over the 2 weeks
# Hydrophilidae = number of hydrophilid beetles (Water Scavenger Beetles)
# Dytiscidae = number of dytiscid beetles (Predaceous Diving Beetles)

# Fish and Nutrient are provided as integers, but I wish to use them as ordered
# factors so that, by default, polynomial contrasts will be used to compute the
# ANOVAs.

df1$Nutrient <- factor(df1$Nutrient,
                       levels = c("0", "4", "8"),
                       ordered = TRUE)

df1$Fish <- factor(df1$Fish,
                   levels = c("0", "1", "2"),
                   ordered = TRUE)

df1$Block <- factor(df1$Block)

```

### Statistical Methods

A linear model was conducted in order to be able to determine the effects of nutrients, fish predators, and combined predators and nutrients on the invertebrate species described previously. A package was used [@lme4] to make a linear mixed-effects model.

```         
```

```{r}
#| label: tbl-Culex
#| tbl-cap: "Linear mixed-effects model of nutrients and predation on Culex egg rafts using data provided by @Pintar.etal2018 "
#| tbl-subcap:
#|     - "Fixed effect."
#|     - "Random effect."
#| layout-nrows: 2

culexmod <- lmer(Culex_Egg_Rafts ~ Fish*Nutrient + (1|Block), data = df1)

CulexAnova <- anova(culexmod)


kbl(
  CulexAnova,
  col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "P"
    ),
  digits = c(3,3,0,0,3,4)
  ) |>
  kable_classic()

kbl(
  rand(culexmod),
  col.names = c(
    "Parameters",
    "log-likelihood",
    "AIC",
    "LRT",
    "df",
    "P"
    ),
  digits = c(0,3,3,3,0,4)
  )
```

```{r}
#| label: tbl-Hydro
#| tbl-cap: "Linears mixed-effect models of nutrients and predation on hydrophilid beetles using data provided by @Pintar.etal2018"
#| tbl-subcap:
#|     - "Fixed effect."
#|     - "Random effect."
#| layout-nrows: 2

hydromod <- lmer(Hydrophilidae ~ Fish*Nutrient + (1|Block), data = df1)

HydroAnova <- anova(hydromod)


kbl(
  HydroAnova,
  col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "P"
    ),
  digits = c(3,3,0,0,3,4)
  ) |>
  kable_classic()

kbl(
  rand(hydromod),
  col.names = c(
    "Parameters",
    "log-likelihood",
    "AIC",
    "LRT",
    "df",
    "P"
    ),
  digits = c(0,3,3,3,0,4)
  )
```

```{r}
#| label: tbl-Dytis
#| tbl-cap: "Linear mixed-effects models of nutrients and predation on dytiscid beetles using data provided by @Pintar.etal2018"
#| tbl-subcap:
#|     - "Fixed effect."
#|     - "Random effect."
#| layout-nrows: 2

dytismod <- lmer(Dytiscidae ~ Fish*Nutrient + (1|Block), df1)

DytisAnova <- anova(dytismod)


kbl(
  DytisAnova,
  col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "P"
    ),
  digits = c(3,3,0,0,3,4)
  ) |>
  kable_classic()

kbl(
  rand(dytismod),
  col.names = c(
    "Parameters",
    "log-likelihood",
    "AIC",
    "LRT",
    "df",
    "P"
    ),
  digits = c(0,3,3,3,0,4)
  )
```

**Results**

There is no significant interaction between fish/predator presence for any Culex, hydrophilid, or dytiscid as shown in @tbl-Culex, @tbl-Hydro, and @tbl-Dytis.

When looking at @tbl-Culex there was a significant difference within the Culex egg rafts whenever there were increasing nutrients (P= 0.0426), and decreasing fish predators being present (P= 0.0016). The block did show a significant difference.

Whenever looking at @tbl-Hydro there appears to be no significant difference on hydrophilid beetles whenever there were less fish present (P= 0.0606) and there seems to be no significant difference when there were more nutrients present (P= 0.1088). The block, however, did show a significant difference.

Whenever looking at @tbl-Dytis there was a significant difference in dytiscid beetles whenever less fish were present (P= 0), as well as whenever more nutrients were present (P = 0.0256). The block did show a significant difference.

## References
